---
title: "**Sudoku Puzzles**"
author: " Arun Kumar Jaganathan, Karthick Pandi, Lakshya Gazaresen, Michael David Gaughan, Poojanjali Vasu, Shubham Rajesh Khamitkar"
output: 
  html_document:   
    code_download: yes
    code_folding: hide
    fig_height: 6
    fig_width: 9
  word_document: default
runtime: shiny

---
<style type="text/css">

body, td {
   font-size: 12px;
   text-align: justify;
}
code.r{
  font-size: 10px;
}
pre {
  font-size: 10px
}
h4.author { 
    font-size: 14px;
  font-family: "Times New Roman", Times, serif;
  color: Black;
  text-align: center;
}
h1.title { 
    font-size: 20px;
  font-family: "Times New Roman", Times, serif;
  color: Black;
  text-align: center;
}
</style>

# {.tabset .tabset-fade .tabset-pills}

## __Introduction__

### {.tabset .tabset-fade}


><font size="2">
<p>The objective of this group statistical report is to analyse the experimental Sudoku Puzzles dataset collected from an experiment conducted in Maynooth University compiled from 2009 to 2013.The dataset contains four types of Sudoku puzzles and the time associated to complete it. A correctly completed Sudoku puzzle will contain all six characters in each row, each column and each box of six cells. Sudoku puzzles are typically a 9 x 9 grid but here mini Sudoku puzzles, with a 6 x 6 grid, are used so that the puzzles can be completed in a short timeframe. The types of Sudoku puzzles differ only in the character required to be filled into each grid.</p>
The types are:
<p>1) Numbers <br />
2) Letters <br />
3) Greek Letters<br />
4) Symbols <br />
</p>
<p>Three hypotheses tests has been carried out based on the dataset. The first hypotheses is to check how the type of Sudoku effects in completing the Sudoku successfully. The second hypotheses is to check how having previous experience of playing Sudoku impacts in successful completion of the puzzles. The third hypotheses is to check if there is any statistical difference in median survival time between the different levels of Sudoku using Survivial Analysis.A Visualisation Dashboard has been created in order to interpret the analysis visually using R Shiny and Tableau.</p>



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## __Data Cleaning and Manipulation__ 

### {.tabset .tabset-fade}

#### __Data Cleaning__ 
```{r}
suppressMessages(library(tidyverse))
```

<p>The 3SudokuCombined.csv dataset has been imported to the current working directory and the dataset contains columns:<br />
1) Class - Indicating class group from 1 to 8<br />
2) Before1 - Have played Sudoku before: yes or no<br />
3) Type - Type of Sudoku puzzle: Numbers, Greek, Letters or Symbols<br />
4) Correct - Puzzle was correct: yes or no<br />
5) Time1 - Time to completion: Mins:Secs<br />
6) Mins - Time to completion: minutes component<br />
7) Seconds - Time to completion: seconds component<br />
8) Time2 - Time to completion: total seconds<br />
9) Before2 - Have played Sudoku before: no, yes in last three months (Within 3 or Inside), yes but not in last three months (Out 3 or Outside)<br />
10) Logic - Enjoy playing logic puzzles: yes, no, indifferent<br />  </p>
<p>A glimps of the dataset can be shown below:</p>

```{r}
sudoku<-read.csv("3SudokuCombined.csv", header=T, na.strings=c("","NA"),stringsAsFactors = F)

sudoku<-as_tibble(sudoku)


########### Data Cleaning (#before,type,correct,sec )###############################

head(sudoku)
```

Columns "Class", "Correct" and "Type" have been changed to factors as they are categorical values and Time2 has been changed to numeric. From the dataset it is observed that there are some ambiguity regarding the spelling of "Types". For instance, some of the observation were recorded as "letter" instead of " Letters", "Number" instead of "Numbers" and "Symbol" instead of "Symbols. In order to avoid ambiguity, the types have been modified respectively. For the "Before2" column, "inside" and "outside" are changed to  "Within3" and "Out_3" respectively as these are basically same.

```{r}
# Class to factor

sudoku$Class<-as.factor(sudoku$Class)

# Response=Correct to factor

sudoku$Correct<-factor(ifelse(sudoku$Correct=='Yes',1,0))


# Time2=total time in sec (change it to numeric)

sudoku$Time2<-suppressWarnings(as.numeric(sudoku$Time2))
# Type to factor


sudoku$Type<-ifelse(sudoku$Type=="Letter","Letters",sudoku$Type)
sudoku$Type<-ifelse(sudoku$Type=="Number","Numbers",sudoku$Type)
sudoku$Type<-ifelse(sudoku$Type=="Symbol","Symbols",sudoku$Type)

sudoku$Type<-factor(sudoku$Type)
#2 Cleaning before2 i.e. inside3 and within_3 are same and out_3 and ouside
# are same

sudoku$Before2<-ifelse(sudoku$Before2=="Inside","Within_3",sudoku$Before2)
sudoku$Before2<-ifelse(sudoku$Before2=="Outside","Out_3",sudoku$Before2)

```

#### __Data Manipulation__
<p>For the data manipulation part, we are filling the missing values of "Before1" column using the values in "Before2" column. If for instance, a player has not played Sudoku within or outside 3 months then definetly the player has not played Sudoku, hence making "Before1" column as No for the player and vice versa. Further if the value of "Before1" column is Yes, it has been classified into 1 or else 0 which gives us the information that there are 388 players who does not have experience in solving Sudoku Puzzles where as 1305 are having experience. </p>

```{r}
## Filling missing values of before1 using before2 details

sudoku%>%
  mutate(Before1=replace(Before1,
                         is.na(Before1)&!is.na(Before2)&Before2=='No',
                         'No')
  )->sudoku     

sudoku%>%
  mutate(Before1=replace(Before1,
                         is.na(Before1)&!is.na(Before2)&Before2!='No',
                         'Yes')
  )->sudoku 

sudoku$Before1<-factor(ifelse(sudoku$Before1=='Yes',1,0))

table(sudoku$Before1)
```

<p> As the dataset contains  Time1(Min:Sec), Mins and Seconds columns, giving us the information about the time taken to finish the Sudoku Puzzles, which are also in consolidated form in Time2 column, we have checked if its possible to impute missing values in Time2 column from these columns, but we couldn't impute any values and hence we dropped Mins, Seconds and Time1 columns.

</p>
```{r}
#### Checking if missing null values in Time2 can be filled using Time1:Min:Sec

check_time<-sudoku%>%filter(is.na(Time2),!is.na(Time1),!is.na(Mins),!is.na(Seconds))

# As it returns zero rows so Time1,Mins and Seconds are not required columns so dropping it.

# sudoku%>%select(-Time1,-Mins,-Seconds)->sudoku

sudoku%>%dplyr::select(-Time1,-Mins,-Seconds)->sudoku



#checking if there is any column filled with value "."

l <- lapply(sudoku,function(x)which((x==".")==T))

sudoku_missing<-sudoku[c(1302,1304,1499,1533,1548),]

# We cannot impute anything for this . 



# Trying to impute before2 using before1 , setting before2='No' if before1=0 

sudoku%>%
  mutate(Before2=replace(Before2,
                         is.na(Before2)&!is.na(Before1)&Before1==0,
                         'No')
  )->sudoku
# 
# sudoku%>%filter(is.na(Before2)&!is.na(Before1)&Before1==0)
```
<p>Further missing values is checked in all the columns and it was observed that there are 26% of data missing in "Before2" column and 34.86% in "Logic" column, hence these columns has been dropped. After cleaning and manipulating the columns of the dataset, the preprocessing of the data is completed and we start with the modelling with the mentioned columns below:</p>

```{r}

perentage_before2<- sudoku%>%filter(is.na(Before2))%>%nrow *100 / nrow(sudoku)


# Since around 26% of data is missing even after imputing, 
# so we skip this column 

percentage_logic<-sudoku%>%filter(is.na(Logic))%>%nrow *100 / nrow(sudoku)


# Since around 34.86% of data is missing and we cannot impute this column,
# so we skip this column 


sudoku%>%dplyr::select(-6,-7)->sudoku

head(sudoku)
#sudoku%>%filter(is.na(Time2))%>%nrow *100 / nrow(sudoku)


# Since around 6% of data is missing so we KEEP this column 

```


## __Hypotheses Testing and EDA__
### {.tabset .tabset-fade}
#### __Dependency Test__
<p>In order to begin with our analysis, we first started with variable analysis, inter dependencies and summarised each types of Sudoku with their minimum, maximum, mean, median ,standard deviations of the completion of the puzzles. For each type of puzzles the number of successful and unsuccessful observations has been tabulated. Also the impact of previous experience and class with respect to correct solution of the puzzle is shown below</p>


```{r}
# Variable Analysis

sudoku%>%filter(Correct==1)%>%
  group_by(Type)%>%
  summarise(min=fivenum(Time2,na.rm = T)[1],
            Q1=fivenum(Time2,na.rm = T)[2],
            median=fivenum(Time2,na.rm = T)[3],
            Q3=fivenum(Time2,na.rm = T)[4],
            max=fivenum(Time2,na.rm = T)[5],
            mean=mean(Time2,na.rm = T),
            iQR=IQR(Time2,na.rm = T),
            SD=sd(Time2,na.rm = T))

table(Type=sudoku$Type,Correct=sudoku$Correct) 
table(Previous_Exp=sudoku$Before1,Correct=sudoku$Correct)
table(Previous_Exp=sudoku$Class,Correct=sudoku$Correct)
```
<p>To find the dependecies between Type, Before1 and Class with the Correct column, which means we are finding how these variables are depending in the correct solution of the Sudoku, we have done Chi Square Test of Independence. It was observed that Type, Before1 and Class are dependent in correct solution of the Sudoku puzzles as the p-values for each of these tests were below 0.05.</p>

```{r}
####### Chi sq test of independence

# correct vs type 
# Chi Sq test of independence to check if there is any dependency between 
# if sudoku was correct and sudoku type at 0.05 significant level



chisq.test(sudoku$Type,sudoku$Correct)

# pvalue<0.05, so there is evidence to prove that there exists dependency
# between correct and sudoku type, This can be noticable
# as we see that greek type has the lowest correct solutions

# correct vs experience
# Chi Sq test of independence to check if there is any dependency between 
# if sudoku was correct and sudoku experience at 0.05 significant level



chisq.test(sudoku$Before1,sudoku$Correct)

# pvalue<0.05, so there is evidence to prove that there exists dependency
# between correct and before hand experience, This can be noticable
# as we see that those who played before got correct solutions more.


# correct vs class
# Chi Sq test of independence to check if there is any dependency between 
# if sudoku was correct and class at 0.05 significant level




chisq.test(sudoku$Class,sudoku$Correct)


# pvalue<0.05, so there is evidence to prove that there exists dependency
# between correct and sudoku class, This can be noticable
# as we see that those who played before got correct solutions more.
```

 Welch Two Sample t-test
```{r}
################################################################################
##T test (only on correct data i.e. Sudoku is correct)


correct_sudoku<-sudoku%>%filter(Correct==1)


sudoku_known<-correct_sudoku%>%filter(Type!='Greek')
sudoku_unknown<-correct_sudoku%>%filter(Type=='Greek')


t.test(sudoku_known$Time2,sudoku_unknown$Time2)


# A Welch Two Sample t-test detected a significant difference(p<0.05) in mean time to succesfully complete
# between known letter and uunknown letter in sudoku(greek).

remove(sudoku_known)
remove(sudoku_unknown)


################################################################################
### One-way ANOVA (only on correct data i.e. Sudoku is correct)

fit<-aov(data=correct_sudoku,Time2~Type)

anova(fit)

# This states there was evidence of an effect of Sudoku type on the
# length of time to Sudoku completion

# TukeyHSD(fit)

remove(fit)
remove(correct_sudoku)
```
#### __Hypotheses__

```{r}
# library(epiDisplay)  

# Logistic Regression on all variable except Time2 as we will do different test for Time2

sudoku[-5]->sudoku_1

na.omit(sudoku_1)->sudoku_1


# Model Building and validations

before  <-glm(Correct~Before1,data = sudoku_1,family="binomial")
type    <-glm(Correct~Type    ,data = sudoku_1,family="binomial")
class   <-glm(Correct~Class,data = sudoku_1,family="binomial")

before_type  <-glm(Correct~Before1+Type,data = sudoku_1,family="binomial")
type_class    <-glm(Correct~Type+Class,data = sudoku_1,family="binomial")
class_before  <-glm(Correct~Class+Before1,data = sudoku_1,family="binomial")

full<-glm(Correct~Before1+Type+Class,data = sudoku_1,family="binomial")

intersect_model<-glm(Correct~Before1*Type*Class,data = sudoku_1,family="binomial")

intercept<-glm(Correct~1,data = sudoku_1,family="binomial")

# Model Bulding and variable significance testing

# Hypothesis 1 : 
# H0 : Intersection is not significant (Reduced model is better)
# H1 : Intersection is significant (full model is better)

anova(full,intersect_model, test = "LR")


# This states Intersection is not significant and thus go with a reduced model(full)
# -----------------------------------------------------------------

# Hypothesis 2 : 
# H0 : Class is not significant (Reduced model is better)
# H1 : Class is significant (full model is better)

anova(before_type,full, test = "LR")


# This states class is not significant and thus go with a reduced model
# -----------------------------------------------------------------
# Hypothesis 3 : 
# H0 : Type is not significant (Reduced model is better)
# H1 : Type is significant (full model is better)

anova(before,before_type, test = "LRT")

# This states Type is significant and thus go with a full model

# ------------------------------------------------------------------
# Hypothesis 4 : 
# H0 : Before1  is not significant (Reduced model is better)
# H1 : Before1  is significant (full model is better)

anova(type,before_type, test = "LRT")

# This states Before1 is significant and thus go with a full model
# ---------------------------------------------------------------------


#Interpretation  
summary(before_type)
exp(before_type$coefficients)

# 
# Hypothesis 1:
#   
#   The odds of successfully completing the Sudoku correctly were 9.81 times higher for 
# those who had played Sudoku before relative to those who had not played before
# 
# Hypothesis 2:
#   
#   The odds of successfully completing the Sudoku correctly were 1.79 times higher for 
# those who had played Sudoku of type "Letters" rather than "Greek",
# The odds of successfully completing the Sudoku correctly were 1.99 times higher for 
# those who had played Sudoku of type "Numbers" rather than "Greek",
# The odds of successfully completing the Sudoku correctly were 1.79 times higher for 
# those who had played Sudoku of type "Symbols" rather than "Greek",
# 
# Hypothesis 3:
#   
#   The class had no effect on successfully completing the sudoku puzzle so as 
# hypothesis 3 we will do survival analysis on Time2





# Approach 2 of variable selection

require(MuMIn) 
model.sel(intercept,before,type,class,before_type,type_class,class_before,full,
          intersect_model) 


# This states as approach1 go with model having Before1 and Type  as it has lowest AIC.


# Approach 3 using StepAIC

library(MASS)
stepAIC(intersect_model,trace = T)$anova

# This also states go with model having Before1 and Type  as it has lowest AIC.



remove(before)
remove(type)
remove(class)
remove(type_class)
remove(class_before)
remove(full)
remove(intercept)
remove(intersect_model)
remove(sudoku_1)

# Accuracy of model is around 81%
table(sudoku$Correct,as.numeric(predict(before_type,newdata = sudoku
                                        ,type = "response")>0.5))


remove(before_type)
```


```{r,eval=FALSE}
############# Survival Analysis on Time2

library(survival)
# install.packages("survminer")
library(survminer)
library(ggfortify)

# Event: 1-> sucesfuly completed
# 0-> not completed
# 
# Time2 is the survival time


sudoku[c(1,2,3,4,5)]%>%as.data.frame->sudoku_survival


fct_explicit_na(sudoku_survival$Before1)->sudoku_survival$Before1

as.character(sudoku_survival$Correct)%>%as.numeric->sudoku_survival$Correct

surv_object <- Surv(time = sudoku_survival$Time2,event = sudoku_survival$Correct)


fit1 <- survfit(surv_object ~ 1, data = sudoku_survival)

fit2 <- survfit(surv_object ~ Type, data = sudoku_survival)
# summary(fit2)


# res.sum <- surv_summary(fit2)

# res<-fortify(fit1)

# # Understanding Summary,
# time is nothing but time2 in ascending order
# n.risk is the the total no of person at risk of dying i.e those who didnt solved puzzle yet.
# n.event is the no of people that solved the event at time=time2
# survival is the no of people still alive in percentage

# https://rpkgs.datanovia.com/survminer/index.html

ggsurvplot(fit1, data = sudoku_survival, pval = T,conf.int = T,
           risk.table = F,censor.size=4,censor.shape="|")

ggsurvplot(fit2, data = sudoku_survival, pval = T,conf.int = T,
           risk.table = F,censor.size=7,censor.shape="|")

# 
# ggplot(data = res, aes(x = time, y = surv, color = strata)) +
#   geom_line() + 
#   # plot censor marks
#   geom_point(aes(shape = factor(ifelse(n.censor >= 1, 1, NA)))) + 
#   # format censor shape as "+"
#   scale_shape_manual(values = 3) + 
#   # hide censor legend 
#   guides(shape = "none") +ggtitle("Kaplan-Meier curves")


# ----------------------------------------

# Time entries on worksheets with incorrect solutions contain information on the 
# quantity of interest since they give LOWER bounds for true completion times.
# The continuous variable time is therefore referred to as right-censored.
# 
# In our context below terms indicates,
#   a. Survival -> Puzzle has not yet correctly completed.
#   b. Death    -> Puzzle has correcly completed.
#   c. Censor point -> TIme of a person who completed puzzle incorrectly.

# T -> Time of death(after correctly completing)
# 
# survival function given as, S(t) = P(T≥ t) for time t.
# 
# The survival function is a function that gives the probability that a patient,
# device, or other object of interest will survive beyond any specified time(here t).
# 
# Using this method, plot shows the estimated survival probabilities(y-axis)
# versus time (x-axis) for each Sudoku type. 
# Censored data points are indicated by a | symbol and the probability of
# completing the Sudoku puzzle correctly by time t is estimated by 1- E(S(t)) 
# (indicates estimated).

fit2

# From the fit1 function we can see correctly completing sudoku given the sudoku type
# is 125 seconds for Numbers, 149 seconds for Letters, 205 seconds for Symbols,
# while 246 seconds for Greek, these unbiased median
# estimates are all higher than the biased medians presented for
# the subset of correct Sudokus only
# 
# Hypothesis:
#   
#   Ho: There is no statistical difference in median survival time between 4 different
# types of sudoku puzzle
# 
# HA: This is statistical difference in median survival time between 4 different
# types of sudoku puzzle



library(coin)

logrank_test(surv_object~Type, data=sudoku_survival)


# From p value we can say accept alternate hypothesis and say 
# There is statistical difference in median survival time between 4 different
# types of sudoku puzzle
```


## __Sudoko Visualisation__ 

### {.tabset .tabset-fade}

#### __Visualisation__ 



```{r}
library(shiny) 

ui <- fluidPage(uiOutput("Link"))

                      server <- function(input, output, session) {

                        output$Link <- renderUI({

                          a(image="tab.png","Sudoku Dashboard", href="https://public.tableau.com/profile/jak94#!/vizhome/SudokuProjectTableauWorkbook/Viz1_1", target="_blank")

                        })}

                      shinyApp(ui = ui, server = server)


```

